ArKanjo: a tool for detecting function-level Code Duplication in the Linux Kernel

### 2025-06-19 paper/paper.tex
Update paper: several adjustments and fix to compile with pdflatex command

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{array}
\usepackage{url}
\usepackage{todonotes}

\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  %\graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage{graphicx}
  \graphicspath{{figuras/},{fig},{logos/},{img/},{images/},{imagens/}}
  % declare the path(s) where your graphic files are
  % \graphicspath{{./eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps, .png, .jpeg}
\fi

@@

The detection of code duplication, or code clones, has been a research subject for 
decades~\cite{firstman}. The literature provides a standard taxonomy, classifying 
clones into four types based on their degree of similarity, from identical copies 
(Type-1) to semantically equivalent but textually different fragments (Type-4)~\cite{litreview}. 
Various detection methodologies have been developed, including textual, token-based, 
tree-based, and graph-based approaches~\cite{litreview}. These have culminated in 
state-of-the-art techniques, such as the graph-based work by Liu et al.~\cite{tailor}. 	
However, the primary focus of such academic work remains on determining whether a given 
pair of code artifacts are duplicates rather than providing a scalable method to scan 
an entire codebase for actionable results. Conversely, existing free software tools explored 
in practice often lack this sophistication, relying instead on primitive textual matching 
that is insufficient for the complexity and scale of the kernel.

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{diagrama_mestrado}
\caption{Architecture diagram with the tool components}
\label{fig:diagram}
\end{figure}

The tool operates in two main parts, as illustrates in Figure~\ref{fig:diagram}: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} performs heavy computations to find code 
duplications across a codebase and produces artifacts to store duplication-related information 
in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This design allows the 
tool to perform the heavy and time-consuming steps only once per codebase, enabling fast performance 
for multiple queries on the same codebase.

The Preprocessor contains two main components: the Function Breaker and the Code Duplication Finder. 
The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase,
extracts the functions along with metadata, and creates a temporary codebase where each extracted 
function becomes a new code file. The Code Duplication Finder then iterates over every pair of files 
in the temporary codebase, checks whether they are code clones, and, if so, stores the result in 
the Code Duplication Database. This database is a text file that records each code duplication as a triple 
\textbf{<function1, function2, similarity>}, where \textbf{function1} and \textbf{function2} are the 
duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
used in the Code Duplication Finder.

The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
information about duplicated functions based on the user's request. If the user executes it
without previously executing the Preprocessor, the Query Responder automatically 
calls the Preprocessor to generate the required artifacts.

The Function Parser receives the input codebase and transforms it into the temporary
codebase. It iterates through every source code file written in a supported programming 
language and uses a language-specific extractor to isolate each function. For each extracted 
function, two new source code files and a metadata file are created in the temporary codebase, 
represented by the pair \textbf{<file name, function name>}, the source code of the function, 
and the function metadata. One of the new files contains the function body, and the other 
contains the function signature. The metadata file includes additional relevant information,
such as the function name, the line where the signature starts, and the line where the function 
body ends. Currently, the supported programming languages are C and Java, although Java support is limited.

@@

@@

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{explorer_example}
\caption{Example of the \textit{Duplication Explorer} functionality.}
\label{fig:explorer_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{function_example}
\caption{Example of the \textit{Function Information} functionality.}
\label{fig:function_ex}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{relatory_example}
\caption{Example of the \textit{Duplication Report} functionality.}
\label{fig:relatory_ex}
\end{figure}

@@

@@

To validate the capabilities of the ArKanjo tool in finding code duplications using the 
text similarity detection method, we applied two independent
evaluation approaches. The first method assesses the tool using a literature-based approach,
comparing it against the BigCloneBench dataset~\cite{bigclonebench}. The second
method consists of an empirical analysis of a subset of functions within the AMD Display driver
that our tool identified as duplicates.

To validate the tool against the BigCloneBench dataset~\cite{bigclonebench}, we followed
the methodology presented by Liu et al.~\cite{tailor}. We sampled 20{,}000 pairs from each clone
type and added 20{,}000 non-duplicate pairs as negative samples. We applied the same sampling
strategy to our tool to ensure a fair comparison.
Unlike state-of-the-art methods, our tool does not distinguish between clone types and
identifies duplications at the function level rather than the file level, which is typical in
state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation
purposes. Specifically, we considered every pair of functions with a similarity metric
equal to or greater than a threshold $X$ as duplicates and marked the corresponding file
pairs. A correctly identified duplication pair is counted as accurate for its clone type, while
an incorrectly inferred pair is considered incorrect across all clone types.
To understand the impact of varying the similarity threshold, we evaluated our tool using
different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%. We then analyzed
and discussed the results.

@@

To assess whether the duplications found by the ArKanjo tool were actionable, this research conducted 
a multi-part ethnographic study. The author and 12 student groups from the
Free Software Development course at the University of São Paulo acted as new contributors 
to the Linux kernel. 
They used the tool to identify duplications within the Industrial I/O (IIO) subsystem and the AMD Display 
driver subsystem refactored the code to address the issues, and submitted their proposed 
fixes as patches to the community maintainers. The students documented 
their refactoring approaches and experiences interacting with the kernel community throughout this process.

@@

\begin{table}[!t]

@@

A discrepancy emerges when detecting non-duplicate pairs between
the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative
samples inferred as duplications), whereas the empirical analysis revealed a higher percentage
of false positives. We propose two potential explanations for this discrepancy.

The first explanation concerns the limitations and known issues with BigCloneBench, as
highlighted by Krinke and Ragkhitwetsagul \cite{bigfail}. The
second explanation relates to the nature of the AMD Display driver, where code artifacts
often share semantic meaning. In contrast, BigCloneBench comprises self-contained code
artifacts with minimal or no shared semantics. Since our tool relies on a text-based code clone
detection approach, it may naturally perform worse on the AMD Display driver than on the
BigCloneBench dataset.

@@

Using the ArKanjo tool, the main author and 12 student groups identified and refactored duplicated 
code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
as patches to the community maintainers. This effort resulted in 13 patch submissions: 
One was from the main author to the AMD Display driver, and 12 were from the student groups, with 11 targeting 
the IIO subsystem and one aimed at the AMD Display driver. The author's patch was successfully accepted and merged into the kernel after a 
lengthy review process, removing over 
500 lines of duplicated code.

@@

\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,referencias}

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{array}
\usepackage{url}
\usepackage{graphicx}
\usepackage{todonotes}

@@

The detection of code duplication, or code clones, has been a research subject for 
decades~\cite{firstman}. The literature provides a standard taxonomy, classifying 
clones into four types based on their degree of similarity, from identical copies 
(Type-1) to semantically equivalent but textually different fragments (Type-4)~\cite{litreview}. 
Various detection methodologies have been developed, including textual, token-based, 
tree-based, and graph-based approaches~\cite{litreview}. These have culminated in 
state-of-the-art techniques, such as the graph-based work by Liu et al.~\cite{tailor}.

However, the primary focus of such academic work remains on determining whether a given 
pair of code artifacts are duplicates rather than providing a scalable method to scan 
an entire codebase for actionable results. Conversely, existing free software tools explored 
in practice, often lack this sophistication, relying instead on primitive textual matching 
that is insufficient for the complexity and scale of the kernel.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=3.5in]{fig/diagrama_mestrado.png}
\caption{Architecture diagram with the tool components}
\label{fig:diagram}
\end{figure}

The tool operates in two main parts, as illustrated in Figure~\ref{fig:diagram}: the \textbf{Preprocessor} and the \textbf{Query Responder}. 
The \textbf{Preprocessor} performs heavy computations to find code 
duplications across a codebase and produces artifacts to store duplication-related information 
in a structured form. The \textbf{Query Responder} consumes the artifacts produced 
by the Preprocessor to execute the tool functionalities as requested. This design allows the 
tool to perform the heavy and time-consuming steps only once per codebase, enabling fast performance 
for multiple queries on the same codebase.

The Preprocessor contains two main components: the Function Breaker and the Code Duplication Finder. 
The workflow of the Preprocessor is as follows: the Function Parser receives the input codebase,
extracts the functions along with metadata, and creates a temporary codebase where each extracted 
function becomes a new code file. The Code Duplication Finder then iterates over every pair of files 
in the temporary codebase, checks whether they are code clones, and, if so, stores the result in 
the Code Duplication Database. This database is a text file that records each code duplication as a triple 
\textbf{$<$function1, function2, similarity$>$}, where \textbf{function1} and \textbf{function2} are the 
duplicated functions, and \textbf{similarity} is the score returned by the duplication detection method 
used in the Code Duplication Finder.

The Query Responder uses the temporary codebase and the Code Duplication Database to extract 
information about duplicated functions based on the user's request. If the user executes the Query Responder
without previously executing the Preprocessor, the Query Responder automatically 
calls the Preprocessor to generate the required artifacts.

The Function Parser receives the input codebase and transforms it into the temporary
codebase. It iterates through every source code file written in a supported programming 
language and uses a language-specific extractor to isolate each function. For each extracted 
function, two new source code files and a metadata file are created in the temporary codebase, 
represented by the pair \textbf{$<$file name, function name$>$}, the source code of the function, 
and the function metadata. One of the new files contains the function body, and the other 
contains the function signature. The metadata file includes additional relevant information,
such as the function name, the line where the signature starts, and the line where the function 
body ends. Currently, the supported programming languages are C and Java, although Java support is limited.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=3.5in]{fig/explorer_example.png}
\caption{Example of the \textit{Duplication Explorer} functionality.}
\label{fig:explorer_ex}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics[width=3.5in]{fig/function_example.png}
\caption{Example of the \textit{Function Information} functionality.}
\label{fig:function_ex}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics[width=4in]{fig/relatory_example.png}
\caption{Example of the \textit{Duplication Report} functionality.}
\label{fig:relatory_ex}
\end{figure}

@@

We applied two independent evaluation approaches to validate the capabilities of the ArKanjo tool in finding code duplications using the text similarity detection method. The first method assesses the tool using a literature-based approach,
comparing it against the BigCloneBench dataset~\cite{bigclonebench}. The second
method consists of an empirical analysis of a subset of functions within the AMD Display driver
that our tool identified as duplicates.

To validate the tool against the BigCloneBench dataset~\cite{bigclonebench}, we followed
the methodology presented by Liu et al.~\cite{tailor}. We sampled 20{,}000 pairs from each clone
type and added 20{,}000 non-duplicate pairs as negative samples. We applied the same sampling
strategy to our tool to ensure a fair comparison.

Unlike state-of-the-art methods, our tool does not distinguish between clone types and
identifies duplications at the function level rather than the file level, which is typical in
state-of-the-art tools. Therefore, we adapted the metric calculation method for evaluation
purposes. Specifically, we considered every pair of functions with a similarity metric
equal to or greater than a threshold $X$ as duplicates and marked the corresponding file
pairs. A correctly identified duplication pair is counted as accurate for its clone type, while
an incorrectly inferred pair is considered incorrect across all clone types.
To understand the impact of varying the similarity threshold, we evaluated our tool using
different threshold values: 30\%, 40\%, 50\%, 60\%, 70\%, 80\%, 90\%, and 100\%. We then analyzed
and discussed the results.

@@

We conducted a multi-part ethnographic study to assess whether the duplications found by the ArKanjo tool were actionable. In the second semester of 2024 and the first semester of 2025, respectively, the main author, conducting a first validation, and 12 student groups (19 undergraduate and 6 graduate students) from the Free Software Development course at the University of São Paulo acted as new contributors to the Linux kernel in a broader experiment, used the tool to identify duplications within the Industrial I/O (IIO) subsystem and the AMD Display driver subsystem refactored the code to address the issues, and submitted their proposed fixes as patches to the community maintainers. The students documented their refactoring approaches and experiences interacting with the kernel community throughout this process.

@@

\begin{table}[ht]

@@

Nevertheless, a discrepancy emerges when detecting non-duplicate pairs between
the two methods. In the BigCloneBench evaluation, we found no false positives (i.e., negative
samples inferred as duplications), whereas the empirical analysis revealed a higher percentage
of false positives. We propose two potential explanations for this discrepancy.

The first explanation concerns the limitations and known issues with BigCloneBench, as
highlighted by Krinke and Ragkhitwetsagul~\cite{bigfail}. The
second explanation relates to the nature of the AMD Display driver, where code artifacts
often share semantic meaning. In contrast, BigCloneBench comprises self-contained code
artifacts with minimal or no shared semantics. Since our tool relies on a text-based code clone
detection approach, it may naturally perform worse on the AMD Display driver than on the
BigCloneBench dataset.

@@

Using the ArKanjo tool, the main author and 26 students (divided into 12 groups) identified and refactored duplicated code in the AMD Display driver and the Industrial I/O (IIO) subsystems, submitting their work 
as patches to the community maintainers. This effort resulted in 13 patch submissions: 
One was from the main author to the AMD Display driver, and 12 were from the student groups, with 11 targeting 
the IIO subsystem and one aimed at the AMD Display driver. After a lengthy review process, the main author's patch was successfully accepted and merged into the kernel, removing over 500 lines of duplicated code.

@@

\newpage
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,referencias}
